\begin{thebibliography}{99}

\bibitem{LeCun1998}
LeCun Y. Gradient-based learning applied to document recognition / Y. LeCun, L. Bottou, Y. Bengio, P. Haffner // Proceedings of the IEEE. -- 1998. -- Vol. 86, № 11. -- P. 2278--2324.

\bibitem{Krizhevsky2012}
Krizhevsky A. ImageNet classification with deep convolutional neural networks / A. Krizhevsky, I. Sutskever, G. Hinton // Advances in Neural Information Processing Systems. -- 2012. -- Vol. 25. -- P. 1097--1105.

\bibitem{Simonyan2015}
Simonyan K. Very deep convolutional networks for large-scale image recognition / K. Simonyan, A. Zisserman // International Conference on Learning Representations. -- 2015.

\bibitem{He2016}
He K. Deep residual learning for image recognition / K. He, X. Zhang, S. Ren, J. Sun // Proceedings of the IEEE Conference on Computer Vision and Pattern Recognition. -- 2016. -- P. 770--778.

\bibitem{Shannon1949}
Shannon C.E. Communication in the presence of noise / C.E. Shannon // Proceedings of the IRE. -- 1949. -- Vol. 37, № 1. -- P. 10--21.

\bibitem{Simoncelli1992}
Simoncelli E.P. Shiftable multiscale transforms / E.P. Simoncelli, W.T. Freeman, E.H. Adelson, D.J. Heeger // IEEE Transactions on information theory. -- 1992. -- Vol. 38, № 2. -- P. 587--607.

\bibitem{Azulay2019}
Azulay A. Why do deep convolutional networks generalize so poorly to small image transformations? / A. Azulay, Y. Weiss // Journal of Machine Learning Research. -- 2019. -- Vol. 20, № 184. -- P. 1--25.

\bibitem{Zhang2019}
Zhang R. Making convolutional networks shift-invariant again / R. Zhang // Proceedings of the International Conference on Machine Learning. -- 2019. -- P. 7324--7334.

\bibitem{Chaman2021}
Chaman A. Truly shift-invariant convolutional neural networks / A. Chaman, I. Dokmanić // Proceedings of the IEEE/CVF Conference on Computer Vision and Pattern Recognition. -- 2021. -- P. 3773--3783.

\bibitem{Redmon2016}
Redmon J. You only look once: Unified, real-time object detection / J. Redmon, S. Divvala, R. Girshick, A. Farhadi // Proceedings of the IEEE Conference on Computer Vision and Pattern Recognition. -- 2016. -- P. 779--788.

\bibitem{Joost-Pieter2016}
Stawiaski J. Shift-invariance, cyclically symmetric filters and periodically expanded convolutions / J. Stawiaski // arXiv preprint arXiv:1609.09296. -- 2016.

\bibitem{Zou2020}
Zou W. On the stability of video object detection / W. Zou, S. Zhu, K. Yu, A. Toshev // arXiv preprint arXiv:2008.05732. -- 2020.

\bibitem{Vasconcelos2021}
Vasconcelos C.N. Understanding pooling in convolutional neural networks / C.N. Vasconcelos, B.N. Silva, M. Fairman // Neurocomputing. -- 2021. -- Vol. 466. -- P. 314--326.

\bibitem{Engstrom2019}
Engstrom L. Exploring the landscape of spatial robustness / L. Engstrom, B. Tran, D. Tsipras, L. Schmidt, A. Madry // Proceedings of the International Conference on Machine Learning. -- 2019. -- P. 1802--1811.

\bibitem{Russakovsky2015}
Russakovsky O. ImageNet Large Scale Visual Recognition Challenge / O. Russakovsky, J. Deng, H. Su, J. Krause, S. Satheesh, S. Ma, Z. Huang, A. Karpathy, A. Khosla, M. Bernstein, A.C. Berg, L. Fei-Fei // International Journal of Computer Vision. -- 2015. -- Vol. 115, № 3. -- P. 211--252.

\bibitem{Lin2014}
Lin T.Y. Microsoft COCO: Common objects in context / T.Y. Lin, M. Maire, S. Belongie, J. Hays, P. Perona, D. Ramanan, P. Dollár, C.L. Zitnick // European Conference on Computer Vision. -- 2014. -- P. 740--755.

\bibitem{Schmidt2020}
Schmidt R.M. Anti-aliased CNNs for improved accuracy and robustness in object detection / R.M. Schmidt, J. Lüing, S. Roth // IEEE Winter Conference on Applications of Computer Vision. -- 2020. -- P. 2738--2747.

\bibitem{Liao2020}
Liao Z. Is object detection with frequency domain invariant a illusion? / Z. Liao, H. Qin, Z. Zha, L. Shao // arXiv preprint arXiv:2012.11704. -- 2020.

\bibitem{Hendrycks2019}
Hendrycks D. Benchmarking neural network robustness to common corruptions and perturbations / D. Hendrycks, T. Dietterich // Proceedings of the International Conference on Learning Representations. -- 2019.

\bibitem{Karras2021}
Karras T. Alias-free generative adversarial networks / T. Karras, M. Aittala, S. Laine, E. Härkönen, J. Hellsten, J. Lehtinen, T. Aila // Advances in Neural Information Processing Systems. -- 2021. -- Vol. 34. -- P. 852--863.

\bibitem{Nyquist1928}
Nyquist H. Certain topics in telegraph transmission theory / H. Nyquist // Transactions of the American Institute of Electrical Engineers. -- 1928. -- Vol. 47, № 2. -- P. 617--644.

\bibitem{Cohen2016}
Cohen T.S. Group equivariant convolutional networks / T.S. Cohen, M. Welling // Proceedings of the International Conference on Machine Learning. -- 2016. -- P. 2990--2999.

\bibitem{Selvaraju2017}
Selvaraju R.R. Grad-CAM: Visual explanations from deep networks via gradient-based localization / R.R. Selvaraju, M. Cogswell, A. Das, R. Vedantam, D. Parikh, D. Batra // Proceedings of the IEEE International Conference on Computer Vision. -- 2017. -- P. 618--626.

\bibitem{Sun2020}
Sun H. Shift equivariance in object detection / H. Sun, E. Lakshminarayanan, B. Li, D. Ramanan // arXiv preprint arXiv:2008.05787. -- 2020.

\bibitem{Song2021}
Song X. Rethinking Object Detector Training: Learning from Circular Shift-Based Invariant Predictions / X. Song, Y. Li, J. Lin, X. Zhang, J. Han // arXiv preprint arXiv:2112.12526. -- 2021.

\bibitem{Ruderman1994}
Ruderman D.L. The statistics of natural images / D.L. Ruderman // Network: computation in neural systems. -- 1994. -- Vol. 5, № 4. -- P. 517--548.

\bibitem{Wang2020}
Wang H. Spatial-Frequency Domain Non-local Network for Image Super-Resolution / H. Wang, Y. Ke, Y. Wen, K. Li, L. Ju, M. Yang, S. Liang // arXiv preprint arXiv:2012.15094. -- 2020.

\bibitem{Xiao2020}
Xiao C. Enhancing adversarial robustness via test-time transformation ensembling / C. Xiao, P. Zhong, C. Zheng // Pattern Recognition. -- 2020. -- Vol. 100. -- P. 107107.

\bibitem{Jaderberg2015}
Jaderberg M. Spatial transformer networks / M. Jaderberg, K. Simonyan, A. Zisserman, K. Kavukcuoglu // Advances in neural information processing systems. -- 2015. -- Vol. 28.

\bibitem{Sagonas2016}
Sagonas C. Shifting towards image-invariant face detection / C. Sagonas, Y. Panagakis, A. Leidinger, S. Zafeiriou // Pattern Recognition. -- 2016. -- Vol. 52. -- P. 125--136.

\bibitem{Vasiljevic2016}
Vasiljevic I. Examining the impact of blur on recognition by convolutional networks / I. Vasiljevic, A. Chakrabarti, G. Shakhnarovich // arXiv preprint arXiv:1611.05760. -- 2016.

\bibitem{Zech2018}
Zech J.R. Variable generalization performance of a deep learning model to detect pneumonia in chest radiographs: A cross-sectional study / J.R. Zech, M.A. Badgeley, M. Liu, A.B. Costa, J.J. Titano, E.K. Oermann // PLoS medicine. -- 2018. -- Vol. 15, № 11. -- P. e1002683.

\bibitem{Geirhos2020}
Geirhos R. The partial invariance of deep convolutional networks / R. Geirhos, C.R.M. Temme, J. Rauber, H.H. Schütt, M. Bethge, F.A. Wichmann // arXiv preprint arXiv:2004.11919. -- 2020.

\bibitem{Sun2019}
Sun Y. Test-time augmentation with uncertainty estimation for deep learning-based medical image segmentation / Y. Sun, X. Wang, Z. Tang, T. Tian, P. Han // arXiv preprint arXiv:1906.04884. -- 2019.

\bibitem{Wusatowski2021}
Wusatowski M. On the Influence of Low-level Features in Object Detection / M. Wusatowski, R. Kaczmarczyk, D. Długosz // arXiv preprint arXiv:2103.10663. -- 2021.

\bibitem{Tan2020}
Tan M. EfficientDet: Scalable and efficient object detection / M. Tan, R. Pang, Q.V. Le // Proceedings of the IEEE/CVF Conference on Computer Vision and Pattern Recognition. -- 2020. -- P. 10781--10790.

\bibitem{Ji2019}
Ji X. Does BlurPool really work? / X. Ji, Y. Zeng, S. Shi, S. Yang, S. Xu // arXiv preprint arXiv:1912.11856. -- 2019.

\bibitem{Wang2020b}
Wang J. On the frequency bias of generative models / J. Wang, Y. Jiang, J. Wu, Z. Sun, Z. Qin, H. Zha // Advances in Neural Information Processing Systems. -- 2020. -- Vol. 33. -- P. 15135--15146.

\bibitem{Worrall2019}
Worrall D.E. Deep scale-spaces: Equivariance over scale / D.E. Worrall, M. Welling // Advances in Neural Information Processing Systems. -- 2019. -- Vol. 32.

\bibitem{Weiler2019}
Weiler M. General E(2)-equivariant steerable CNNs / M. Weiler, G. Cesa // Advances in Neural Information Processing Systems. -- 2019. -- Vol. 32.

\bibitem{Ultrych2019}
Ultrych J. On invariance of convolutional models to input padding / J. Ultrych, R. Š. Picek, J. Matas // IEEE Transactions on Pattern Analysis and Machine Intelligence. -- 2019. -- Vol. 41, № 12. -- P. 3020--3026.

\bibitem{Bietti2019}
Bietti A. On the inductive bias of neural tangent kernels / A. Bietti, J. Mairal // Advances in Neural Information Processing Systems. -- 2019. -- Vol. 32.

\bibitem{Bruna2013}
Bruna J. Invariant scattering convolution networks / J. Bruna, S. Mallat // IEEE Transactions on Pattern Analysis and Machine Intelligence. -- 2013. -- Vol. 35, № 8. -- P. 1872--1886.

\end{thebibliography} 